---
output:
  reprex::reprex_document:
    venue: "gh"
    advertise: FALSE
    session_info: TRUE
    style: TRUE
    comment: "#;-)"
    tidyverse_quiet: FALSE
    std_out_err: TRUE
knit: reprex::reprex_render
---

In this notebook, we attempt to compute the food desert index using
* Satellite imagery to identify population distribution
* Kernel density estimate of supermarket locations

Since the method 

We attempt to overlay the two rasters as matrices and evaulate the 
magnitude of the Hadamard product (element-wise multiplication) of the two matrices.


```{r set_up}
knitr::opts_chunk$set(echo = TRUE)


library(stringr)
library(data.table)
library(purrr)
library(dplyr)

library(gstat)

library(sf)
library(spdep)
library(tmap)
library(spatstat)
library(mapview)
library(ggplot2)
library(maptools)

library(raster)
library(terra)
# library(rgdal)
```

## 1. Loading the Data
Since we will be overlaying the rasters onto the satellite image,
which are in the crs of 4326.

```{r load_data}
# load england shp
eng_shp_path <- file.path("data","eng_boundary_buffer.geojson")
eng_shp <- read_sf(eng_shp_path)  %>%
            dplyr::filter(name == "England")
eng_owin <- as.owin(eng_shp) # in epsg 27700

# load msoa shp
msoa_path <- file.path("data","MSOA_2021")
msoa_shp <- read_sf(msoa_path) # in epsg 27700

# load msoa rural urban classification
ru_class <- read.csv("./data/ons_census_2021.csv")  
ru_class <- ru_class  %>% 
          mutate(urb = grepl("Urban",RUC11, fixed = T)) 

# load the supermarkets
spmkt_path <- file.path("data","amenities_supermarket_processed.geojson")
spmkt_shp <- read_sf(spmkt_path)  %>% 
                dplyr::filter(shop == "supermarket")  %>% 
                st_transform(27700)                
# add coordinate
coord <- spmkt_shp %>% 
    st_coordinates()  %>% 
    as.data.frame()  
spmkt_shp <- cbind(spmkt_shp, coord) 
spmkt_shp_ppp <- as.ppp(spmkt_shp)  

# load the raster
satellite_paths <- file.path("satellite")  %>% 
    list.files(full.names = T, pattern = "*.tif")
```

Return a window object from the the list of supermarkets location to compute the kernel density.

```{r win_obj_spmkt}
class(eng_shp)
Window(spmkt_shp_ppp) <- eng_owin
# visualise the location of the supermarkets
mapview(spmkt_shp)
```

## 2. Compute Kernel Densities and Visualise it.
Method 2 Using ggplot
https://ggplot2.tidyverse.org/reference/geom_density_2d.html#aesthetics
In R, I created a kernel density map given a geospatial data. I want to change the way the choropleth map is printed. 
Specifically, I want to bin the data points into 5 equal quantiles. 

```{r viz_kde}
# density map data
spmkt_shp_kernel <- spmkt_shp  %>% 
                st_crop(eng_shp)
# Define a custom function to calculate 5 equal quantiles
quantile_breaks <- function(density_output) {
  quantile(density_output, probs = seq(0, 1, length.out = 5))
}
# define legend labels
custom_labels <- c("Very Low", "Low", "Medium", "High", "Very High")


# create the background map
eng_background <- ggplot()+
  geom_sf(data = eng_shp, inherit.aes = FALSE) +  # plot the shapefile
  xlab("") + ylab("") 
# overlay kernel density map  
eng_map <- eng_background +
          stat_density2d(data=spmkt_shp, # from the ggplot2 package
                  aes(x=X, y=Y, fill=after_stat(level), 
                  alpha=after_stat(level)), 
                  # geom = "raster", contour = F
                  adjust = 1.5, # control bandwidth
                  geom="polygon",
                  ) +
#  scale_alpha_continuous(range=c(0.25,0.65)) +
scale_fill_gradientn(colors = rev(viridis::magma(5)), breaks = quantile_breaks, labels = custom_labels) +
labs(fill = "Supermarket Density") + 
# scale_color_viridis_d(option = "magma", breaks = quantile_breaks) + 
 theme(line = element_blank(),    # customize the plot style
        rect = element_blank(), 
        axis.text=element_blank()) + 
 guides(fill=guide_legend("Supermarket"), alpha = "none") # add legend

eng_map  

# ggsave("./img/kernel_density_spmkt.png", dpi = 300)

```

## 3. Compute the Hadamard Product of Urban Classification and Supermarket KDE 

First, let's test for the pipeline before iterating through all the msoa.

```{r viz_sat_img}
id <- 66
# return the msoa
msoa21cd <- str_extract(satellite_paths[id], "E\\d+")
msoa_bound <- msoa_shp %>% dplyr::filter(MSOA21CD == msoa21cd)
# reproject satellite image
sat_img_temp <- raster(satellite_paths[id]) #in epsg 4326

sat_img_27700 <- projectRaster(sat_img_temp,
                          crs = CRS("+init=epsg:27700"))
# return the raster as a df for plotting
raster_df <- as.data.frame(rasterToPoints(sat_img_27700), xy = TRUE)
# rename the columns
colnames(raster_df) <- c("x","y","value")

raster_df$value <- as.factor(ifelse(is.na(raster_df$value), "NA", as.character(raster_df$value)))
# Create a dummy data frame for the boundary line to be included in the legend
boundary_legend <- data.frame(
  x = NA, y = NA, label = "Boundary"
)

# overlay MSOA boundary over raster object
ggplot() +
  geom_raster(data = raster_df, aes(x = x, y = y, fill = value)) +
  scale_fill_manual(
    values = c("NA" = "#FFFFFF", "1" = "#E3978B"),
    labels = c("NA" = "Non-residential", "1" = "Residential"),
    name = ""
  )+
  # scale_fill_gradientn(colours=c("#000000","#FFFFFF")) +  
  geom_sf(data = msoa_bound, fill = "#cdcdcd55", color ="#8EAAC0", linewidth = 2) +
  geom_sf(data = msoa_bound, aes(color = "Boundary"), fill = NA, linewidth = 1, show.legend = TRUE) +
    scale_color_manual(
    values = c("Boundary" = "#8EAAC0"),
    labels = c("Boundary" = "Bexley 004 MSOA"),
    name = "Legend"
  ) + 
   theme(
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_blank(),
  ) 
ggsave("./img/E02000068_overlay.png")
```

From the output above, we observe that the raster is aligned with the shapefile that st_coordinates
the adminstrative boundary of the MSOAs.

The below chunk requires the input of :
* specified MSOA satellite image `sat_img_27700`
* Boundary of the MSOA `msoa_bound`

```{r est_fd_ind}
# return the extent of the raster
mask <- raster::raster(sat_img_27700)
# rasterize the msoa boundary
mask <- rasterize(msoa_bound, mask, field=1, fun='first')

# Get the matrix representation of the raster and the mask
raster_matrix <- as.matrix(sat_img_27700)
mask_matrix <- as.matrix(mask)

# Perform the Hadamard product
result_matrix <- raster_matrix * mask_matrix

# Create a new raster object from the result matrix
result_raster <- raster(result_matrix)

# Set the extent and CRS of the new raster to match the original raster
extent(result_raster) <- extent(sat_img_27700)
crs(result_raster) <- crs(sat_img_27700)


# show the masked residential area
ggplot() +
  geom_raster(data = result_raster, aes(x = x, y = y, fill = value)) +
  scale_fill_manual(
    values = c("NA" = "#FFFFFF", "1" = "#E3978B"),
    labels = c("NA" = "Non-residential", "1" = "Residential"),
    name = ""
  )+
  # scale_fill_gradientn(colours=c("#000000","#FFFFFF")) +  
  geom_sf(data = msoa_bound, fill = "#cdcdcd55", color ="#8EAAC0", linewidth = 2) +
  geom_sf(data = msoa_bound, aes(color = "Boundary"), fill = NA, linewidth = 1, show.legend = TRUE) +
    scale_color_manual(
    values = c("Boundary" = "#8EAAC0"),
    labels = c("Boundary" = "Bexley 004 MSOA"),
    name = "Legend"
  ) + 
   theme(
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_blank(),
  ) 

```


```{r kde_est}
# Compute the kde of the supermarkets
# return the buffer dimension of the msoa dependent on rural urban classifiation
urbs <- ru_class  %>% dplyr::filter(MSOA21CD == msoa21cd)  %>% dplyr::select("urb")
if (urbs[[1,1]]){
  buffer <- 1000
} else {buffer <- 5000}

# return the relevant subset of supermarket
# Create a 10km buffer around the polygon
buffer_bound <- st_buffer(msoa_bound, dist = buffer + 1)
pois_within_buffer <- spmkt_shp[st_within(spmkt_shp, buffer_bound, sparse = FALSE), ]
# Convert POIs to SpatialPoints
pois_sp <- as(pois_within_buffer, "Spatial")
# Define the window (extent) of the KDE to match the result raster
ext <- extent(result_raster)
win <- as.owin(as(ext, "SpatialPolygons"))

# Convert to ppp object
pois_ppp <- as(pois_sp, "ppp")
pois_ppp$window <- win

# Define the resolution to match the result_raster
res_x <- xres(result_raster)
res_y <- yres(result_raster)

# Perform KDE using spatstat
kde <- density(pois_ppp, sigma = buffer, at = "pixels", eps = c(res_x, res_y))

# Convert KDE result to raster
kde_raster <- raster(kde, crs=crs(result_raster))

if (!all(dim(kde_raster) == dim(result_raster))) {
  stop("The dimensions of the KDE raster and the result raster do not match.")
}

# Convert rasters to matrices
kde_matrix <- as.matrix(kde_raster)

# Perform the Hadamard product
hadamard_matrix <- kde_matrix * result_matrix

# Create a new raster object from the result matrix
hadamard_raster <- raster(hadamard_matrix)

# Set the extent and CRS of the new raster to match the original rasters
extent(hadamard_raster) <- extent(result_raster)
crs(hadamard_raster) <- crs(result_raster)

# Validation visualisation 
# return the raster as a df for plotting
raster_df <- as.data.frame(rasterToPoints(hadamard_raster), xy = TRUE)
a <- max(raster_df$layer)

ggplot() +
  geom_raster(data = raster_df, aes(x = x, y = y, fill = layer)) +
  scale_fill_gradientn(colours=c("#FFFFFF","#048EF8"), name = "Supermarket KDE"
          # limits = c(0, a)
          ) +  
  geom_sf(data = msoa_bound, fill = NA, color = "red", linewidth = 2) +
  geom_sf(data = pois_within_buffer, aes(shape = "Supermarket"), size  =3) + 
   scale_shape_manual(name = "", values = c("Supermarket" = 16)) +
   theme(
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_blank(),
    # legend.position = "none"
  ) 

# -------------------
```

From the visualisation, this seems like a possible method to compute food access.
Let's compute the average value of KDE density within the built up area in the MSOA.
```{r msoa_summary}
dplyr::summarise(spmkt_ind_rast)
mean(spmkt_ind_rast$layer)
```

Let's pack all the codes above into a function
```{r kde_func}
compute_spmkt <- function(file_path){
    # return the msoa code and boundary given the file path
    msoa21cd <- str_extract(file_path, "E\\d+")
    msoa_bound <- msoa_shp %>% dplyr::filter(MSOA21CD == msoa21cd)
    # load and reproject satellite image
    sat_img_temp <- raster(file_path) #in epsg 4326
    sat_img_27700 <- projectRaster(sat_img_temp,
                            crs = CRS("+init=epsg:27700"))
    # return the raster as a df for plotting
    raster_df <- as.data.frame(rasterToPoints(sat_img_27700), xy = TRUE)
    # rename the columns
    colnames(raster_df) <- c("x","y","value")

    # mask the urban classifiaction raster
    # return the extent of the raster
    mask <- raster::raster(sat_img_27700)
    # rasterize the msoa boundary
    mask <- rasterize(msoa_bound, mask, field=1, fun='first')

    # Get the matrix representation of the raster and the mask
    raster_matrix <- as.matrix(sat_img_27700)
    mask_matrix <- as.matrix(mask)

    # Perform the Hadamard product -------------
    result_matrix <- raster_matrix * mask_matrix

    # Create a new raster object from the result matrix
    result_raster <- raster(result_matrix)

    # Set the extent and CRS of the new raster to match the original raster
    extent(result_raster) <- extent(sat_img_27700)
    crs(result_raster) <- crs(sat_img_27700)

    # Compute the kde of the supermarkets ------------------
    # return the buffer dimension of the msoa dependent on rural urban classifiation
    urbs <- ru_class  %>% filter(MSOA21CD == msoa21cd)  %>% dplyr::select("urb")
    if (urbs[[1,1]]){
      buffer <- 1000
    } else {buffer <- 5000}

    # return the relevant subset of supermarket
    # Create a 1/5 km buffer around the polygon
    buffer_bound <- st_buffer(msoa_bound, dist = buffer + 1)
    pois_within_buffer <- spmkt_shp[st_within(spmkt_shp, buffer_bound, sparse = FALSE), ]
    # Convert POIs to SpatialPoints
    pois_sp <- as(pois_within_buffer, "Spatial")
    # Define the window (extent) of the KDE to match the result raster
    ext <- extent(result_raster)
    win <- as.owin(as(ext, "SpatialPolygons"))

    # Convert to ppp object
    pois_ppp <- as(pois_sp, "ppp")
    pois_ppp$window <- win

    # Define the resolution to match the result_raster
    res_x <- xres(result_raster)
    res_y <- yres(result_raster)

    # Perform KDE using spatstat
    kde <- density(pois_ppp, sigma = buffer, at = "pixels", eps = c(res_x, res_y))

    # Convert KDE result to raster
    kde_raster <- raster(kde, crs=crs(result_raster))

    if (!all(dim(kde_raster) == dim(result_raster))) {
      stop("The dimensions of the KDE raster and the result raster do not match.")
    }

    # Convert rasters to matrices
    kde_matrix <- as.matrix(kde_raster)

    # Perform the Hadamard product
    hadamard_matrix <- kde_matrix * result_matrix

    # Create a new raster object from the result matrix and parse as df
    hadamard_raster <- raster(hadamard_matrix)
    raster_df <- as.data.frame(rasterToPoints(hadamard_raster), xy = TRUE)

    # output the value 
    val <- mean(raster_df$layer)
    return(val[[1]])
}
```

### 3.1 Iterate Through the MSOA

First, we need to find the list of MSOAs that satify the following:
* Was not recorded in `./data/spmkt_spec_satellite.csv`
* Satellite image is export

```{r for_loop_kde_est}
# find the list of MSOAs for which, we have the satellite image
# return the list of msoa and their satellite image
# load the raster
satellite_paths <- file.path("satellite")  %>% 
    list.files(full.names = T, pattern = "*.tif")
file_ids <- str_extract(satellite_paths, "E\\d+")

# return the list of msoas recorded
if (file.exists("./data/spmkt_spec_satellite.csv")){
                  msoa_processed <- read.csv("./data/spmkt_spec_satellite.csv") %>%
                  dplyr::select("MSOA21CD")
    } else {msoa_processed <- data.frame(MSOA21CD = NA)}                

# Find exclusive IDs
exclusive_ids <- setdiff(file_ids, msoa_processed$MSOA21CD)

# Subset file paths based on exclusive IDs
exclusive_file_paths <- satellite_paths[file_ids %in% exclusive_ids]

length(exclusive_file_paths)
```


Now, we can iterate through the msoas for which we can process to compute the food desert indicator.

```{r}
#instantiate an empty list to store the result
spmkt_res <- list()
error_list <- c()

for (path in exclusive_file_paths){
  tryCatch(
            {value <- compute_spmkt(path)
            # print(paste0("the value is ", value))
            msoacd21 <- str_extract(path, "E\\d+")
            spmkt_res[[msoacd21]] <- value
            },
    error = function(e){
          print(paste0("An error occured for ", str_extract(path, "E\\d+")))
          error_list <-c(error_list,str_extract(path, "E\\d+")) }
  )
}

df <- tibble::enframe(spmkt_res, name = "MSOA21CD", value = "value")  %>% 
    mutate(value = map_dbl(value, ~ .x))

# export the data
file_path <- file.path(".","data","spmkt_spec_satellite.csv")
if (file.exists(file_path)){
  # specify row names
  id_max <- read.csv(file_path)  %>% dplyr::select(X)  %>% max()
  rownames(df) <- c((id_max + 1): (id_max + dim(df)[1]))
  # append row to existing csv
  write.table(df, file = file_path, sep = ",", col.names = FALSE, append = TRUE)

  } else {write.csv(df, file_path)}

```



### 3.2 Exploratory visualisation

```{r}
# merge the
temp <- msoa_shp  %>%   
  left_join(df, by = "MSOA21CD")  %>% 
  mutate(value = as.numeric(value))

# create plot
ggplot(temp) +
geom_sf(aes(value)) +
scale_fill_viridis_c() + 
theme_void()
```



